Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-07-17 Documents/M1/dgM1Curvilinear.tex
Adding results from test problems.  Text for the Milne problem

To model neutrino matter interactions, we include emission and absorption, (isotropic) elastic scattering on nucleons and nuclei, and neutrino-electron scattering (NES).  
On the right-hand sides of Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, $\tilde{\chi}(\epsilonNu,\bU)$ and $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$ are absorption and total (absorption plus scattering) opacities, respectively.  
The absorption opacity $\tilde{\chi}$ is corrected for stimulated absorption \citep[cf.][]{mezzacappaBruenn_1993a}.  
The NES rates $R^{\In}$ and $R^{\Out}$ are obtained by expanding the full kernels in a Legendre series, retaining only the isotropic part \citep[see, e.g.,][]{cernohorsky_1994}, and the integrals extend over all energies $\bbR^{+}=\{\epsilonNu~|~\epsilonNu\in[0,\infty)\}$.  
The energy volume element is $dV_{\epsilonNu}=\epsilonNu^{2}d\epsilonNu$.  
The opacities depend on the neutrino energy and the local thermodynamic state of the fluid $\bU=(\rho,T,Y_{e})^{T}$, where $\rho$ is the (baryon) mass density, $T$ the temperature, and $Y_{e}$ the electron fraction.  
For simplicity we consider a single neutrino species (electron neutrinos), and we use opacities given by \citet{bruenn_1985}; see Section~\ref{sec:opacities} for further details.  
In Eq.~\eqref{eq:momentEquationEnergy}, the equilibrium density $\cJ_{0}=4\pi\,f_{0}(\epsilonNu,\bU)$ is computed from the angular moment of the isotropic Fermi-Dirac distribution
\begin{equation}
  f_{0}(\epsilonNu,\bU)=\f{1}{e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}+1},
\end{equation}
where $k_{\mbox{\tiny B}}$ is Boltzmann's constant, and $\mu_{\nu}=\mu_{p}-\mu_{n}+\mu_{e}$ is the neutrino chemical potential.  
(The chemical potentials of protons, neutrons, and electrons are denoted by $\mu_{p}$, $\mu_{n}$, and $\mu_{e}$, respectively.)

@@

For the forthcoming description it is useful to write the moment equations in a more compact form.  
To this end, we define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $\cK^{i}_{~j}$, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
(Explicit expressions for metric functions appearing in the geometry sources are listed in Table~\ref{tab:metricFunctions}.)  
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\bcH\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\chi\,(\cJ_{0}-\cJ)+\cL_{\NES},\,-(\kappa+\kappa_{\NES})\,\bcH\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Here, $\cL_{\NES}$ and $\kappa_{\NES}$ represent the integral operators appearing in Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, respectively.  
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

For non-stiff problems, we employ the strong stability-preserving Runge-Kutta (SSP-RK) methods \citep[e.g.,][]{gottlieb_etal_2001}, which, for evolving the moments from $t^{n}$ to $t^{n+1}=t^{n}+\dt$ ($\bcM^{n}\to\bcM^{n+1}$), take the general form
\begin{align}
  \bcM^{(0)}
  &=\bcM^{n}, \\
  \bcM^{(i)}
  &=\sum_{k=1}^{i-1}\alpha_{ik}\,\big[\,\bcM^{(k)}+\beta_{ik}\,\dt\,\bcL(\bcM^{(k)})\,\big], \quad i=1,\ldots,m 
  \label{eq:ssprkStages} \\
  \bcM^{n+1}
  &=\bcM^{(m)}.  
\end{align}
For second- and third-order temporal accuracy (denoted SSP-RK2 and SSP-RK3, respectively), the coefficient matrices $\alpha_{ik}$ and $\beta_{ik}$ are given in Table~\ref{tab:ssprkCoefficients}.  
Note that $\alpha_{ik},\beta_{ik}\ge0$ and $\sum_{k}\alpha_{ik}=1~\forall i$, so that the RK stages in Eq.~\eqref{eq:ssprkStages} are simply convex combinations of forward Euler steps with time step $\beta_{ik}\,\dt$ \citep{shuOsher_1988}.

@@

To overcome this time step restriction, we integrate the collision term with implicit methods.  
Following \citet{chertock_etal_2015}, we define the forward Euler stages in Eq.~\eqref{eq:ssprkStages}
\begin{equation}
  \bcM_{ik}^{\mbox{\tiny FE}}
  =\bcM^{(k)}+\beta_{ik}\,\dt\,\big(\,\bcF_{\bx}(\bcM^{(k)})+\bcC(\bcM^{(k)})\,\big).  
\end{equation}
The semi-implicit Runge-Kutta (SIRK) method is obtained by substituting $\bcM_{ik}^{\mbox{\tiny FE}}\to\bcM_{ik}^{\mbox{\tiny SI}}$, where $\bcM_{ik}^{\mbox{\tiny SI}}$ is obtained by solving the implicit equation
\begin{equation}
  \bcM_{ik}^{\mbox{\tiny SI}}
  =\bcM^{(k)}+\beta_{ik}\,\dt\,\big(\,\bcF_{\bx}(\bcM^{(k)})+\bcC(\bcM_{ik}^{\mbox{\tiny SI}})\,\big).  
\end{equation}
As discussed by \citet{chertock_etal_2015}, the time-integration method using Eq.~\eqref{eq:ssprkStages}, with $\bcM_{ik}^{\mbox{\tiny FE}}$ replaced by $\bcM_{ik}^{\mbox{\tiny SI}}$, is at most first-order accurate in time.  
To increase the temporal accuracy to second order, \citet{chertock_etal_2015} propose a correction step to be applied after the last stage.  
We defer the development of the correction step, and use the first-order accurate time integrator for stiff problems.  
We note that the SIRK method reduces to the SSP-RK methods in Section~\ref{sec:explicitTime} when $\bcC=0$.

@@

During time integration we apply limiters that modify the polynomial representation in elements so as to prevent development and growth of nonphysical oscillations (slope limiters; Sec.~\ref{sec:slopeLimiters}) or nonphysical states (positivity limiters; Sec.~\ref{sec:positivityLimiters}).

@@

After computing the limited slopes in Eq.~\eqref{eq:limitedSlopes}, they are compared with the original slopes.  
Then, if
\begin{equation}
  |\widetilde{\bcC}_{|\vect{1}|}^{i}-\bcC_{|\vect{1}|}^{i}|/\max\big(|\widetilde{\bcC}_{|\vect{1}|}^{i}|,|\bcC_{|\vect{1}|}^{i}|\big)>\text{tol},
\end{equation}
where we set $\text{tol}=10^{-2}$, we replace the modal expansion in Eq.~\eqref{eq:conservedModalExpansion} with the limited polynomial
\begin{equation}
  \bcM_{\DG}(\bx)
  \to\widetilde{\bcM}_{\DG}(\bx)
  =\langle\bcM\rangle_{\bK}+\sum_{i=1}^{d}\widetilde{\bcC}_{|\vect{1}|}^{i}\,\phi_{|\vect{1}|}^{i}(\bx).  
  \label{eq:conservedModalExpansionLimited}
\end{equation}
We limit all the fields in $\bcM$ if the difference in any of the coefficients exceeds the tolerance.  
This limiting procedure locally destroys the high-order accuracy of the DG method, but leads to non-oscillatory evolution when sharp features are present in the solution.

@@

@@

When solving the moment equations, we need to prevent non-physical states from developing.  
In particular, from the definition of the moments in Eq.~\eqref{eq:angularMoments}, it is easy to show that a positive distribution function implies a positive density $\cJ>0$ and a limited flux density $\cJ-|\bcH|>0$.  
Failure in maintaining these bounds can lead to an ill-posed closure problem and loss of hyperbolicity for the moment equations.

@@

\section{Benchmark Problems}

@@

\section{Mini-app: Transport Problems with Tabulated Neutrino Opacities}

@@

\subsubsection{Neutrino-Electron Scattering}

@@

>>>>

To model neutrino matter interactions, we include emission and absorption, (isotropic) elastic scattering on nucleons and nuclei, and neutrino-electron scattering (NES).  
On the right-hand sides of Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, $\tilde{\chi}(\epsilonNu,\bU)$ and $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$ are absorption and transport (absorption plus elastic scattering) opacities, respectively.  
The absorption opacity $\tilde{\chi}$ is corrected for stimulated absorption \citep[cf.][]{mezzacappaBruenn_1993a}.  
The NES rates $R^{\In}$ and $R^{\Out}$ are obtained by expanding the full kernels in a Legendre series, retaining only the isotropic part \citep[see, e.g.,][]{cernohorsky_1994}, and the integrals extend over all energies $\bbR^{+}=\{\epsilonNu~|~\epsilonNu\in[0,\infty)\}$.  
The energy volume element is $dV_{\epsilonNu}=\epsilonNu^{2}d\epsilonNu$.  
The opacities depend on the neutrino energy and the local thermodynamic state of the fluid $\bU=(\rho,T,Y_{e})^{T}$, where $\rho$ is the (baryon) mass density, $T$ the temperature, and $Y_{e}$ the electron fraction.  
For simplicity we consider a single neutrino species (electron neutrinos), and we use opacities given by \citet{bruenn_1985}; see Section~\ref{sec:opacities} for further details.  
In Eq.~\eqref{eq:momentEquationEnergy}, the equilibrium density $\cJ_{0}=4\pi\,f_{0}(\epsilonNu,\bU)$ is computed from the angular moment of the isotropic Fermi-Dirac distribution
\begin{equation}
  f_{0}(\epsilonNu,\bU)=\f{1}{e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}+1},
\end{equation}
where $k_{\mbox{\tiny B}}$ is Boltzmann's constant, and $\mu_{\nu}=\mu_{p}-\mu_{n}+\mu_{e}$ is the neutrino chemical potential.  
(The chemical potentials of protons, neutrons, and electrons are denoted by $\mu_{p}$, $\mu_{n}$, and $\mu_{e}$, respectively.)

@@

For the forthcoming description it is useful to write the moment equations in a more compact form.  
To this end, we define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $\cK^{i}_{~j}$, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
(Explicit expressions for metric functions appearing in the geometry sources are listed in Table~\ref{tab:metricFunctions}.)  
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\bcH\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\tilde{\chi}\,(\cJ_{0}-\cJ)+\cL_{\NES},\,-(\kappa+\kappa_{\NES})\,\bcH\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Here, $\cL_{\NES}$ and $\kappa_{\NES}$ represent the integral operators appearing in Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, respectively.  
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

For non-stiff problems, we employ the strong stability-preserving Runge-Kutta (SSP-RK) methods \citep[e.g.,][]{gottlieb_etal_2001}, which, for evolving the moments from $t^{n}$ to $t^{n+1}=t^{n}+\dt$ ($\bcM^{n}\to\bcM^{n+1}$) using $m$ stages, take the general form
\begin{align}
  \bcM^{(0)}
  &=\bcM^{n}, \\
  \bcM^{(i)}
  &=\sum_{k=0}^{i-1}\alpha_{ik}\,\big[\,\bcM^{(k)}+\beta_{ik}\,\dt\,\bcL(\bcM^{(k)})\,\big], \quad i=1,\ldots,m 
  \label{eq:ssprkStages} \\
  \bcM^{n+1}
  &=\bcM^{(m)}.  
\end{align}
For second- and third-order temporal accuracy (denoted SSP-RK2 and SSP-RK3, respectively), the coefficient matrices $\alpha_{ik}$ and $\beta_{ik}$ are given in Table~\ref{tab:ssprkCoefficients}.  
Note that $\alpha_{ik},\beta_{ik}\ge0$ and $\sum_{k}\alpha_{ik}=1~\forall i$, so that the RK stages in Eq.~\eqref{eq:ssprkStages} are simply convex combinations of forward Euler steps with time step $\beta_{ik}\,\dt$ \citep{shuOsher_1988}.

@@

To overcome this time step restriction, we integrate the collision term with implicit methods.  
Following \citet{chertock_etal_2015}, we define the forward Euler stages in Eq.~\eqref{eq:ssprkStages}
\begin{equation}
  \bcM_{ik}^{\mbox{\tiny FE}}
  =\bcM^{(k)}+\beta_{ik}\,\dt\,\big(\,\bcF_{\bx}(\bcM^{(k)})+\bcC(\bcM^{(k)})\,\big).  
\end{equation}
The semi-implicit Runge-Kutta (SIRK) method is obtained by substituting $\bcM_{ik}^{\mbox{\tiny FE}}\to\bcM_{ik}^{\mbox{\tiny SI}}$, where $\bcM_{ik}^{\mbox{\tiny SI}}$ is obtained by solving the implicit equation
\begin{equation}
  \bcM_{ik}^{\mbox{\tiny SI}}
  =\bcM^{(k)}+\beta_{ik}\,\dt\,\big(\,\bcF_{\bx}(\bcM^{(k)})+\bcC(\bcM_{ik}^{\mbox{\tiny SI}})\,\big).  
\end{equation}
As discussed by \citet{chertock_etal_2015}, the time-integration method using Eq.~\eqref{eq:ssprkStages}, with $\bcM_{ik}^{\mbox{\tiny FE}}$ replaced by $\bcM_{ik}^{\mbox{\tiny SI}}$, is at most first-order accurate in time.  
To increase the temporal accuracy to second order, \citet{chertock_etal_2015} propose a correction step to be applied after the last stage.  
We defer the development of the correction step \citep[see, e.g.][]{endeveHauck_2017}, and use the first-order accurate time integrator for stiff problems.  
We note that the SIRK method reduces to the SSP-RK methods in Section~\ref{sec:explicitTime} when $\bcC=0$.

@@

During time integration we apply limiters that modify the polynomial representation in elements in order to prevent development and growth of nonphysical oscillations (slope limiters; Sec.~\ref{sec:slopeLimiters}), or nonphysical states (positivity limiters; Sec.~\ref{sec:positivityLimiters}).

@@

After computing the limited slopes in Eq.~\eqref{eq:limitedSlopes}, they are compared with the original slopes.  
Then, if
\begin{equation}
  |\widetilde{\bcC}_{|\vect{1}|}^{i}-\bcC_{|\vect{1}|}^{i}|/\max\big(|\widetilde{\bcC}_{|\vect{1}|}^{i}|,|\bcC_{|\vect{1}|}^{i}|\big)>\text{tol},
\end{equation}
where we set $\text{tol}=10^{-2}$, we replace the modal expansion in Eq.~\eqref{eq:conservedModalExpansion} with the limited polynomial
\begin{equation}
  \bcM_{\DG}(\bx)
  \to\widetilde{\bcM}_{\DG}(\bx)
  =\langle\bcM\rangle_{\bK}+\sum_{i=1}^{d}\widetilde{\bcC}_{|\vect{1}|}^{i}\,\phi_{|\vect{1}|}^{i}(\bx).  
  \label{eq:conservedModalExpansionLimited}
\end{equation}
We limit all the fields in $\bcM$ if the difference in any of the coefficients exceeds the tolerance.  
This limiting procedure locally destroys the high-order accuracy of the DG method, but leads to non-oscillatory evolution when sharp features are present in the solution.  
(In our implementation we apply the limiting to characteristic variables, using the eigenvectors given in Appendix~\ref{sec:eigenstructure}, which gives somewhat better results near discontinuities.)

@@

\ee{Discuss discontinuity detection using Harten's sub-cell resolution method.}

@@

When solving the moment equations, we need to prevent non-physical states from developing.  
In particular, from the definition of the moments in Eq.~\eqref{eq:angularMoments}, it is easy to show that a positive distribution function implies a positive density $\cJ>0$ and a limited flux density $\cJ-|\bcH|>0$.  
Failure in maintaining these bounds can lead to an ill-posed closure problem and loss of hyperbolicity for the moment equations.  
In order to maintain these bounds we adopt the framework of \citet{ZS2010b}, developed for the Euler equations of gas dynamics \citep[see also][]{ZS2010a,ZS2011,OHF2012}.  
The full details of the application of this framework to the multi-dimensional two-moment method for radiation transfer will be presented elsewhere \citep{endeveHauck_2017}.  
We give only a brief description here.  
The bound-preserving methodology introduces the realizable set
\begin{equation}
  \cR=\big\{\bcM=\big(\cJ,\bcH\big)^{\mbox{\tiny T}}~|~\cJ\ge\epsilon\quad\text{and}\quad\gamma(\bcM)=\cJ-|\bcH|\ge\epsilon\big\},
  \label{eq:realizableSet}
\end{equation}
where $\epsilon$ is an arbitrarily small positive parameter.  
The realizable set in Eq.~\eqref{eq:realizableSet} is a convex cone \citep{OHF2012}.  
Key to the \emph{realizability-preserving} (RP) scheme is to express the updated cell-averages of the moments as positive combinations of elements in $\cR$.  
Realizability of the cell-average of the updated moments is then guaranteed by convexity arguments.  
Sufficient conditions include: (i) the polynomial representation of the moments are realizable in a finite set of quadrature points in each element $\bK$, and (ii) the timestep $\dt$ is restricted by a CFL-like condition, which is somewhat stricter than that required for stability.

Since the RP scheme only guarantees realizability of the cell-averages, polynomial limiting is needed after each timestep (and after each stage in the Runge-Kutta method) to enforce realizability in the quadrature points.  
This limiting procedure is similar to that described by \citet{ZS2010a} for the Euler equations \citep[see also][]{liuOsher_1996}, and consists of two steps.  
First, if $\cJ<\epsilon$ for any of the quadrature points, we replace $\cJ_{\DG}\to\widehat{\cJ}_{\DG}$, where the new polynomial representation is given by
\begin{equation}
  \widehat{\cJ}_{\DG}=\theta_{1}\,\cJ_{\DG}+(1-\theta_{1})\,\cJ_{\bK},
\end{equation}
where
\begin{equation}
  \theta_{1}=\min\Big\{1,\f{\cJ_{\bK}-\epsilon}{\cJ_{\bK}-\cJ_{\mbox{\tiny min}}}\Big\},
\end{equation}
and $\cJ_{\mbox{\tiny min}}$ is the minimum density in all the quadrature points.  
(Note that the limiting procedure is conservative, since it leaves the cell average unchanged.)  
In the second step, we let $\widehat{\bcM}_{\DG}=(\widehat{\cJ}_{\DG},\bcH_{\DG})^{\mbox{\tiny T}}$.  
Then, if $\widehat{\bcM}_{\DG}$ lies outside $\cR$ for any quadrature point (i.e., $\gamma(\widehat{\bcM}_{\DG})<\epsilon$), there exists an intersection point of the straight line connecting $\bcM_{\bK}\in\cR$ and $\widehat{\bcM}_{\DG}$ evaluated in the troubled quadrature point (denoted $\widehat{\bcM}_{q}$), and the surface of $\cR$.  
Any point on this line is given by the convex combination $\bs_{q}(\xi)=(1-\xi)\,\bcM_{\bK}+\xi\,\widehat{\bcM}_{q}$, where $\xi\in[0,1]$, and the intersection point is obtained by solving the equation $\gamma(\bs_{q}(\xi))=\epsilon$ for $\xi$.  
We then replace the polynomial representation $\widehat{\bcM}_{\DG}\to\widetilde{\bcM}_{\DG}$, where
\begin{equation}
  \widetilde{\bcM}_{\DG}=\theta_{2}\,\widehat{\bcM}_{\DG}+(1-\theta_{2})\,\bcM_{\DG},
\end{equation}
and $\theta_{2}$ is the smallest $\xi$ obtained in the element by considering all the troubled quadrature points.

@@

\section{Basic Benchmark Problems}

@@

The Milne problem \citep[see, e.g.,][for details]{hummerRybicki_1971} considers isotropic radiation emanating from a source with radius $r=R_{\mbox{\tiny min}}$, propagating through a purely (elastic) scattering atmosphere with an outer boundary at $r=R_{\mbox{\tiny max}}$.  
Thus, the absorption opacity is zero $\tilde{\chi}=0$, and the scattering opacity is given by a power law in radius
\begin{equation}
  \sigma=r^{-n}.  
  \label{eq:opacityMilne}
\end{equation}
\citet{hummerRybicki_1971} computed steady-state solutions for $n=3/2$, $2$, and $3$, and different values of $R_{\mbox{\tiny max}}$ by solving the radiative transfer equation.  
Here we compare results obtained with the two-moment model with solutions given by \citet{hummerRybicki_1971}.

With the scattering opacity in Eq.~\eqref{eq:opacityMilne}, the optical depth at radius $r$ is given by
\begin{equation}
  \tau(r)=\f{1}{n-1}\big[\,1/r^{n-1}-1/R_{\mbox{\tiny max}}^{n-1}\,\big].
\end{equation}
\citet{hummerRybicki_1971} also provide asymptotic solutions for the energy density in the optically thick and thin regimes, which we use to set initial values and boundary conditions when solving the two-moment equations.  
In the optically thick regime (large $\tau$), the asymptotic value for the energy density is
\begin{equation}
  \cJ_{\mbox{\tiny thick}}(r)
  =\f{3}{r^{2}}\,\Big(\,\f{n-1}{n+1}\,\Big)\,[\,\tau+K\,], 
  \label{eq:MilneThick}
\end{equation}
where $K$ is a constant.  
In the optically thin regime (small $\tau$), the asymptotic value is
\begin{equation}
  \cJ_{\mbox{\tiny thin}}(r)
  =\f{1}{r^{2}}\,[\,1+\tau\,].  
\end{equation}
Our initial condition for the energy density is then set as
\begin{equation}
  \cJ_{0}(r)
  =(1-e^{-\tau(r)})\,\cJ_{\mbox{\tiny thick}}(r)+e^{-\tau(r)}\,\cJ_{\mbox{\tiny thin}}(r),
\end{equation}
while the initial momentum density is set as $\cH_{r,0}(r)=1/r^{2}$.  
These expressions are also used to set boundary conditions, which are held fixed during the computations.  
(We use solutions provided by \citet{hummerRybicki_1971} in the thick regime to determine values for the constant $K$ in Eq.~\eqref{eq:MilneThick}.)
With the specified initial and boundary conditions, we evolve the moment equations until a steady state is reached.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.4]{./Figures/MilneProblem1D_n20.png} &
      \includegraphics[scale=0.4]{./Figures/MilneProblem1D_n20_EddingtonFactors.png} \\
      \includegraphics[scale=0.4]{./Figures/MilneProblem1D_n15.png} &
      \includegraphics[scale=0.4]{./Figures/MilneProblem1D_n30.png}
    \end{tabular}
  \end{center}
  \caption{Results of the Milne problem obtained with the two-moment model (solid lines), obtained with the DG(2)+SIRK2 scheme, compared with solutions provided by \citep{hummerRybicki_1971} (boxes).  See text for details.}
  \label{fig:milneProblem1D}
\end{figure}

Results for various combinations of $n$ and $R_{\mbox{\tiny max}}$, obtained with the DG2+SIRK2 scheme, are plotted in Figure~\ref{fig:milneProblem1D}.  
The radius of the inner boundary is $R_{\mbox{\tiny min}}=0.01$ in all the models.  
We use 64 elements for the models with $R_{\mbox{\tiny max}}=0.1$, and 128 elements for the models with $R_{\mbox{\tiny max}}=10$, $30$, and $100$.  
The size of the innermost (smallest) element is $\Delta r=10^{-3}$, which increases geometrically with radius.

In the upper left panel in Figure~\ref{fig:milneProblem1D}, we plot $\cJ\times r^{2}$ versus radius for $n=2$, where $R_{\mbox{\tiny max}}=0.1$ (solid black lines) and $R_{\mbox{\tiny max}}=30$ (solid blue lines).  
These results are compared with solutions given by \citet{hummerRybicki_1971} (black and blue boxes).  
We also plot the asymptotic values in the thick regime (dashed lines), while the radius where $\tau=1$ is indicated with dotted vertical lines.  
The the two lower panels show similar results using different values of $n$ and $R_{\mbox{\tiny max}}$.  
In the lower left panel we plot results for $n=3/2$, with $R_{\mbox{\tiny max}}=0.1$ (black lines) and $R_{\mbox{\tiny max}}=100$.  
In the lower right panel we plot results for $n=3$, with $R_{\mbox{\tiny max}}=0.1$ (black lines) and $R_{\mbox{\tiny max}}=10$.  
In the upper right panel we plot flux factors ($\cH_{r}/\cJ$, black lines) and Eddington factors ($\psi$, blue lines) for a model with $n=2$ and $R_{\mbox{\tiny max}}=10$, and compare with results given \citet{hummerRybicki_1971} (their Table~I; boxes).  
For this model, we computed results with the Minerbo closure (cf. Eq.~\eqref{eq:eddingtonFactor}; solid lines) and the Levermore closure (cf. Eq.~\eqref{eq:eddingtonFactorLevermore}; dashed lines).  
As can be seen, there is very little difference between the closure methods, but the results obtained with the Levermore closure are somewhat closer to the results given by \citet{hummerRybicki_1971}.

Overall, there is quite good agreement between the results obtained with the two-moment model and the transport solutions of \citet{hummerRybicki_1971} when $R_{\mbox{\tiny max}}/R_{\mbox{\tiny min}}=10$.  
For larger values of this ratio, the agreement is worse, but seems to improve with increasing value of the index $n$ (cf. the blue lines in the lower two panels in Figure~\ref{fig:milneProblem1D}).

\section{Transport Problems with Tabulated Neutrino Opacities}

@@

\subsubsection{Neutrino-Electron Scattering (NES)}

When NES is included the internal energy of the fluid evolves according to
\begin{align}
  \rho\,d_{t}\epsilon
  &=-\int_{\bbR^{+}}\int_{\bbR^{+}}\big(4\pi-\cJ(\epsilonNu)\big)R^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}\,\epsilonNu\,dV_{\epsilonNu} \\
  &\hspace{12pt}
  +\int_{\bbR^{+}}\int_{\bbR^{+}}\cJ(\epsilonNu)\,R^{\Out}(\epsilonNu,\epsilonNu')\,\big(4\pi-\cJ(\epsilonNu)\big)\,dV_{\epsilonNu'}\,\epsilonNu\,dV_{\epsilonNu}.  
\end{align}
With the semi-implicit approximation (evaluating the NES scattering rates using fluid properties at $t^{n}$), the neutrino number density can be solved for without iterative coupling to the thermodynamic state of the fluid.  
We first define the solution vector $\bcJ=(\cJ_{1},\ldots,\cJ_{M^{\epsilonNu}})^{\mbox{\tiny T}}$, and solve the nonlinear system of equations
\begin{equation}
  \bF(\bcJ^{n+1})=0
  \label{eq:NonlinearSystemNES}
\end{equation}
using Newton's method.  
Here, the $j$th component of the nonlinear system of equations is given by
\begin{equation}
  F_{j}(\bcJ^{n+1})
  =\big(\cJ_{j}^{n+1}-\cJ_{j}^{n}\big)
  -\dt\,\Big\{\,
    \big(4\pi-\cJ_{j}^{n+1}\big)\sum_{k=1}^{M^{\epsilonNu}}W_{k}^{(2)}\,R_{jk}^{\In,n}\,\cJ_{k}^{n+1}
    -\cJ_{j}^{n+1}\sum_{k=1}^{M^{\epsilonNu}}W_{k}^{(2)}\,R_{jk}^{\Out,n}\,\big(4\pi-\cJ_{k}^{n+1}\big)\,\Big\},
\end{equation}
and the Jacobian matrix is
\begin{align}
  \pderiv{\bF}{\bcJ}
  =\pderiv{F_{j}}{\cJ_{k}}
  &=\delta_{jk}\,\Big(1+\dt\,\Big\{\sum_{l=1}^{M^{\epsilonNu}}W_{l}^{(2)}\,R_{jl}^{\In,n}\,\cJ_{l}
  +\sum_{l=1}^{M^{\epsilonNu}}W_{l}^{(2)}\,R_{jl}^{\Out,n}\,\big(4\pi-\cJ_{l}\big)\Big\}\Big) \nonumber \\
  &\hspace{12pt}
  -\dt\,\Big\{\big(4\pi-\cJ_{j}\big)\,W_{k}^{(2)}\,R_{jk}^{\In,n}+\cJ_{j}\,W_{k}^{(2)}\,R_{jk}^{\Out,n}\Big\}.
\end{align}
When Eq.~\eqref{eq:NonlinearSystemNES} is converged to give $\bcJ^{n+1}$, the fluid energy density is updated as
\begin{align}
  \epsilon^{n+1}
  &=\epsilon^{n}-\f{\dt}{\rho}
  \Big\{\,
    \sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(3)}\big(4\pi-\cJ_{j}^{n+1}\big)
    \sum_{k=1}^{M^{\epsilonNu}}W_{k}^{(2)}\,R_{jk}^{\In,n}\,\cJ_{k}^{n+1} \nonumber \\
    &\hspace{56pt}
    -\sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(3)}\,\cJ_{j}^{n+1}
    \sum_{k=1}^{M^{\epsilonNu}}W^{(2)}\,R_{jk}^{\Out,n}\,\big(4\pi-\cJ_{k}^{n+1}\big)
  \,\Big\}.
\end{align}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.75\textwidth]{./Figures/RelaxationEmissionAbsorption_Spectra_ModelA}
  \end{center}
  \caption{Results from solving an emission/absorption relaxation problem with initial material properties given by Model A in Table~\ref{tab:tabulatedModels}.  
  The neutrino energy spectrum is plotted for various times as the radiation field approaches equilibrium.}
  \label{fig:RelaxationEmissionAbsorption}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.85\textwidth]{./Figures/RelaxationNES_Spectra.png}
    \begin{tabular}{cc}
      \includegraphics[width=0.475\textwidth]{./Figures/RelaxationNES_DensitiesVsTime.png} &
      \includegraphics[width=0.475\textwidth]{./Figures/RelaxationNES_EnergiesVsTime.png}
    \end{tabular}
  \end{center}
  \caption{Results from solving an NES relaxation problem with initial material properties given by Model C in Table~\ref{tab:tabulatedModels}.}
  \label{fig:RelaxationNES}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[width=0.475\textwidth]{./Figures/RelaxationNES_DensitiesVsTime_Timestep.png} &
      \includegraphics[width=0.475\textwidth]{./Figures/RelaxationNES_EnergiesVsTime_Timestep.png}
    \end{tabular}
  \end{center}
  \caption{Results from the NES relaxation problem with initial material properties given by Model C in Table~\ref{tab:tabulatedModels}, where the timestep $\dt$ has been set to $10^{-5}$~ms (solid lines), $10^{-4}$~ms (dashed lines), and $10^{-3}$~ms (dotted lines).  In the left panel we plot the occupation density versus time for $\epsilonNu=11.2$~MeV (red lines) and $\epsilonNu=97.5$~MeV (blue lines).  In the right panel we plot the fluid energy versus time.}
  \label{fig:RelaxationNES_Timestep}
\end{figure}
